---
title: "R Notebook"
output: html_notebook
---

# Week 6, day 3 - Homework

## Question 1

```{r}
library(tidyverse)
phone_data <- read_csv("data/20190928-items.csv") %>% 
  janitor::clean_names()
```

```{r}
skimr::skim(phone_data)
```

## Question 2
Find the brand with the highest number of phones
```{r}
phone_data %>% 
  group_by(brand) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))
```
Samsung have the highest number of phones. This is significantly higher than any other brand.

## Question 3
For your top brand, plot the distribution of phone ratings as a probability density, overlaying a fitted normal distribution. Do you think the normal distribution provides a good model of these ratings?
```{r}
phone_stats <- phone_data %>% 
  filter(brand == "Samsung") %>%
  summarise(
    mean = mean(rating),
    sd = sd(rating)
  )


phone_data %>% 
  filter(brand == "Samsung") %>% 
  select(rating) %>% 
  ggplot(aes(rating)) +
  geom_histogram(aes(y = after_stat(density)),
                 colour = "white") +
  stat_function(
    fun = dnorm,
    args = list(
      mean = phone_stats$mean,
      sd = phone_stats$sd
    ),
    colour = "blue"
  )
```
Observations

* The data is slightly left skewed, with most ratings around 3-4.
* The normal distribution fits the general shape of the data reasonably well, although:
  - The line is quite a bit higher than most of the histogram bars
  - There are quite a few bars that go significantly higher than the Normal line. This is likely the cause of the line shift up.
  - There are no ratings greater than 5 which leaves a blank area at that side of the normal distribution line.

## Question 4
We can be more quantitative in comparing the distribution of top brand ratings with a normal distribution. You calculated the mean() and sd() of your distribution in an earlier question. Use these values to calculate the proportions of ratings within one-, two- and three standard deviations of the mean. Compare these proportions with those you expect for a normal distribution.

```{r}
samsung_data <- phone_data %>% 
  filter(brand == "Samsung") %>% 
  select(rating)

mean_sam <- phone_stats$mean
sd_sam <- phone_stats$sd

# samsung_data %>% 
#   mutate(within_sd = case_when(
#     rating >= mean_sam - sd_sam & rating <= mean_sam + sd_sam ~ "1",
#     rating >= mean_sam - (2*sd_sam) & rating <= mean_sam + (2*sd_sam) ~ "2",
#     rating >= mean_sam - (3*sd_sam) & rating <= mean_sam + (3*sd_sam) ~ "3",
#     .default = ">3"
#   )) %>% 
#   group_by(within_sd) %>% 
#   summarise(total = n()) %>% 
#   mutate(prop = total/nrow(samsung_data))

sigma_1 <- samsung_data %>% 
  filter(rating >= mean_sam - sd_sam & rating <= mean_sam + sd_sam) %>% 
  summarise(prop = n()/nrow(samsung_data))

sigma_2 <- samsung_data %>% 
  filter(rating >= mean_sam - (2*sd_sam) & rating <= mean_sam + (2*sd_sam)) %>% 
  summarise(prop = n()/nrow(samsung_data))

sigma_3 <- samsung_data %>% 
  filter(rating >= mean_sam - (3*sd_sam) & rating <= mean_sam + (3*sd_sam)) %>% 
  summarise(prop = n()/nrow(samsung_data))

sigma_1
sigma_2
sigma_3
```
Found prop    | Expected prop
--------------|---------------
sigma 1 = 79% |  68%
--------------|---------------
sigma 2 = 91% |  95%
--------------|---------------
sigma 3 = 97% |  99.7%

* The proportions found in the data do not match a normal distribution particularly well.
* The data is slightly more spread out from the mean than a normal distribution.

