---
title: "2_dplyr_practise_questions"
output: html_notebook
---

```{r}
library(CodeClanData)
library(tidyverse)
```

## Question 1. Load in the data, and have a look at it. How many variables does it have? How many observations? What are the variable names? What type of variables does it have?

```{r}
glimpse(state_income_data)
head(state_income_data)
dim(state_income_data)
names(state_income_data)
```


## Question 2.Select the columns which contain the information from years 2002 to 2008. Ensure you keep the Index and State variables as well.
```{r}
income_data_till_2008 <- select(state_income_data, 1:9)
income_data_till_2008
```


```{r}
a_to_c_income_data <- income_data_till_2008 %>% 
  rename(first_letter = Index) %>% # Question 3.Rename the Index column to first_letter. Hint: look up the rename() function.
  filter(first_letter %in% c("A", "B", "C")) # Question 4. Find the seven states which start with an A,B, or C.
a_to_c_income_data


```

```{r}
income_data_change <- a_to_c_income_data %>% 
  arrange(desc(Y2008)) %>%  # Question 5. Find which of these seven states have the greatest income in 2008. = Alabama
  mutate(income_change = Y2008 - Y2002) # Question 6. Calculate the change in income for each state of these seven states (states starting with A,B, or C) between 2002 and 2008. Save the income change in a new column called income_change.
  
filter(income_data_with_change, income_change > 500000 | income_change < -500000) # Question 7. Find which of these seven states income change is greater than 500,000 (i.e. an increase of 500,000 or decrease of 500,000). = Alabama, Colorado

summarise(income_data_change, mean(income_change), median(income_change))# Question 8. Calculate the mean and median income change for all seven states.
```

## Question 9. Go back to the original dataset which contains all states. Write a pipe which selects the State and Y2008 variables, arranges the income from highest to lowest, filters all those states which have a income of greater than 1 million, and then finally chooses the top 10 wealthiest states.

```{r}
state_income_data %>% 
  select(State, Y2008) %>% 
  arrange(desc(Y2008)) %>% 
  filter(Y2008 > 1000000) %>% 
  head(10)
```

