---
title: "R Notebook"
output: html_notebook
---
# Week 5, day 2, Advanced ggplot homework

## Question 1.
Load in the libraries you need, and look at the backpack data. Take note of variable names and dimensions.
```{r}
library(tidyverse)
library(CodeClanData)
```
```{r}
backpack
names(backpack)
dim(backpack)
```

## Question 2. What is the distribution of backpack weights that students are carrying? Choose an appropriate graph to show the distribution.
```{r}
backpack %>% 
  ggplot(aes(backpack_weight)) +
  geom_histogram(binwidth = 1, colour = "white") +
  scale_x_continuous(breaks = seq(0, 40, 5)) +
  theme_classic()
```
This plot illustrates the distribution of backpack weights among University Students.  
While the range is 2-35 kg the most common backpack weight is 10 kg.   
In general the weights are between 5-20 kg.

## Question 3. Are there differences in backpack weight between sex?
```{r}
# backpack %>% 
#   ggplot(aes(backpack_weight, fill = sex)) +
#   geom_bar(position = "dodge") +
#   facet_wrap(~ sex, ncol = 1) +
#   scale_y_continuous(breaks = seq(0, 10, 2)) +
#   theme_light()
```
```{r}
backpack %>% 
  #group_by(sex) %>% 
  #summarise(average_weight = mean(backpack_weight)) %>% 
  ggplot(aes(sex, backpack_weight, fill = sex)) +
  geom_boxplot(show.legend = FALSE) +
  scale_y_continuous(breaks = seq(0,40,2)) +
  theme_classic()
```
On average, males have heavier backpacks by around 1.5 kg. While there is a single outlier at 35 kg, the distribution does confirm that males tend to have heavier backpacks than females.

## Question 4.What is the relationship between body weight and backpack weight?
```{r}
backpack %>% 
  ggplot(aes(body_weight, backpack_weight)) +
  geom_point() +
  geom_smooth(method = "lm")
```
While backpack weight appears to slightly increase with body weight this is not a clear correlation.

## Question 5.Is this relationship different between men and women? Choose an appropriate plot and interpret your findings.
```{r}
backpack %>% 
  ggplot(aes(body_weight, backpack_weight, colour = sex)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  facet_wrap(~ sex)
```
The increase in backpack weight with body weight is slightly clearer in women than in men.

## Question 6. What proportion of students report back pain?
```{r}
backpack %>% 
  mutate(back_problems = if_else(back_problems == 1,
                                  "yes",
                                  "no")) %>% 
  group_by(back_problems) %>% 
  arrange(desc(back_problems)) %>% 
  summarise(proportion = paste0(n()/nrow(backpack)*100, "%")) %>% 
  ggplot(aes(x = "", y = proportion, fill = as.factor(back_problems))) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  coord_polar("y", start=0) +
  labs(x = " ",
       y = " ",
       title = "Proportion of students with back pain") +
  geom_text(aes(label = back_problems))
```
???? Labels do not work and this is a complicated way of doing a simple calculation.
32% of students suffer from back pain.

## Question 7. Is this proportion different between men and women?
```{r}
backpack %>% 
  mutate(back_problems = if_else(back_problems == 1,
                                  "yes",
                                  "no")) %>% 
  ggplot(aes(sex, fill = as.factor(back_problems))) +
  geom_bar(position = "dodge")
```
A much higher proportion of female students experienced back pain compared to male students.

## Question 8. Now you’ve investigated your data separately, create plots to describe the relationship between backpack weight, body weight, gender and reported backpain. Try both backpack_weight and body_weight as well as the ratio.
```{r}

```


# 3 Extensions
## Question 1.

Does reported backpain differ across the year groups? Ignore year 0 in your data.

## Question 2.

Is there a difference between how heavy the bags those in different years carry?