---
title: "R Notebook"
output: html_notebook
---

# Week 2, day 2 - Homework

## Question 1.
Load the code_clan_tweets.csv data. Find the number of rows, columns, and list all the variable names.
```{r}
library(tidyverse)

codeclan_twitter <- read_csv("data/code_clan_tweets.csv")
```
```{r}
dim(codeclan_twitter)
names(codeclan_twitter)
```
## Question 2.
Find the total number of favourites (stored in favorite_count) that CodeClan tweets have got. Don’t add a tweet’s favorite_count to the total when the tweet was quoted (see the is_quote variable).
```{r}
# codeclan_twitter %>% 
#   summarise(missing = sum(is.na(favorite_count)))

codeclan_twitter %>% 
  filter(is_quote == FALSE) %>% 
  summarise(total_faves = sum(favorite_count))
```
## Question 3.
Summarise the mean number of retweets (retweet_count), split by the type of platform that was used to tweet (source). Again, omit quoted tweets from the mean.
```{r}
codeclan_twitter %>% 
  filter(is_quote == FALSE) %>% 
  group_by(source) %>% 
  summarise(mean_retweets = mean(retweet_count))
```

## Question 4.
Count the total number of likes (i.e. total of favorite_count), split by media type, and arrange them from most likes to least. Assume that any tweets without a listed media type are of type “text”.
```{r}
codeclan_twitter %>% 
  mutate(media_type = coalesce(media_type, "text")) %>% 
  group_by(media_type) %>% 
  summarise(total_likes = sum(favorite_count)) %>% 
  arrange(desc(total_likes))
```

## Question 5.
Find the mean number of characters that a CodeClan tweet contains. You can either find a way to count the text in the text variable, or use the variable display_text_width which contains a count. Checking both would be a good idea if possible to see if they match (but it’s not essential).
```{r}
codeclan_twitter %>% 
  summarise(my_mean_chr_count = mean(nchar(text)),
            their_mean_chr_count = mean(display_text_width))
#These give different values, why??
```

```{r}
#TESTS
test_string_1 <- "@CodeClanScot has arrived in the Highlands! https://t.co/IJT6zbhEBT via @YouTube" #table gives 100
str_length(test_string_1)
nchar(test_string_1) #these give 80
#nchar and str_length return the same value. This is different than the display_text_width in the table even when using the same strings. 
#are the symbols confusing these?
test_string_2 <- "@C! https://t.co/IT @Y" #there are 22 characters here
str_length(test_string_2)
nchar(test_string_2) #these return 22

#perhaps the table is counting differently??
codeclan_twitter %>% 
  select(text, display_text_width) %>% 
  mutate(character_count = nchar(text))
```

## Question 6.
The code_clan_info.csv data file contains status_url along with other info. Load this in, and join it to the code_clan_tweets tibble, so that you have a status_url for each tweet. Decide which variable to join the tibbles on.
```{r}
codeclan_info <- read_csv("data/code_clan_info.csv")
```
```{r}
codeclan_joined <- left_join(codeclan_twitter, codeclan_info, by = "tweet_id")
codeclan_joined
```

## Question 7.
From your new joined data, create a new tibble codeclan_hashtags containing only tweet_id and hashtags, then convert hashtags into lowercase for analysis. Keep only those tweets with hashtags.
```{r}
codeclan_hashtags <- codeclan_joined %>% 
  select(tweet_id, hashtags) %>% 
  filter(!is.na(hashtags)) %>% 
  mutate(hashtags = str_to_lower(hashtags))
codeclan_hashtags
```

4 Extensions


## Question 8.
Some tweets have more than one hashtag, denoted by the c( symbols at the start of the string. Use the str_sub() and str_detect() functions and your codeclan_hashtags tibble from above to find all the cases in which the hashtag string begins with charactersc(.
```{r}
codeclan_hashtags %>% 
  filter(str_detect(hashtags, "^c\\("))
```



## Question 9.
Use the str_detect() function to find all tweets with text that mentions “Edinburgh”, and count how many such tweets there are.
```{r}
codeclan_twitter %>% 
  filter(str_detect(text, "Edinburgh")) %>% 
  summarise(total = n())
```


## Question 10.
Use str_extract_all() with a regular expression to find out which Twitter users CodeClan have been tweeting.
```{r}
unique(flatten_chr(str_extract_all(codeclan_twitter$text, "@[:alnum:]+")))

# I also tried with a pipe but kept getting errors ??
codeclan_twitter %>% 
  str_extract_all(text, "@[:alnum:]+")
```

