---
title: "Week 1 - weekend homework"
output: html_notebook
---

# Weekend Homework

First steps  
Load necessary packages and read in books.csv. Investigate dimensions, variables, missing values - you know the drill!

Up to you  
Now it’s up to you… For this weekend homework there will be no specific tasks, just you and this dataset! Using everything you’ve learned this week, try to describe/summarise at least 5 things about this dataset - using R and the tidyverse of course! Feel free to find and use new functions if there is something that the tidyverse doesn’t offer, but do try to use this homework to apply what you have learned this week. Be prepared to share one of your findings on Monday!

Remember  
Before you submit, go through your weekend homework and make sure your code is following best practices as laid out in the coding_best_practice lesson.

## Step 1: Investigate the data
```{r}
library(tidyverse)

books <- read_csv("data/books.csv")
```

```{r}
glimpse(books)
View(books)
```

```{r}
books %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

There are no "NA" missing values

```{r}
books_short <- books %>% #removed identifiers such as ISBN
  select(title:average_rating, language_code:publisher) 

books_short %>% 
  arrange(average_rating)
  
books_short %>% 
  arrange(ratings_count)
```

Some books have a rating figure despite having a total ratings count of 0. Where are these ratings coming from??

Investigate pages
```{r}
books_short %>% 
  arrange(num_pages)
```

Some books have 0 pages. This does not seem likely. '0's should be replaced with NAs if we wish to analyse book length.

Instructors code:
```{r}
books_short %>% 
  select_if(is.numeric) %>% 
  psych::describe(fast = TRUE)
```

### Clean up steps
Set no pages or no ratings to NA entries. Remove "NOT A BOOK" entries.
```{r}
books_clean <- books_short %>% 
  mutate(average_rating = if_else(ratings_count == 0, NA, average_rating),
        num_pages = if_else(num_pages == 0, NA, num_pages)) %>% 
  filter(authors != "NOT A BOOK") #completely remove the entries which are not actually books

books_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

## Step 2 - Analysis

### QUESTION 1: Most and least popular books

The min average_ratings all also have 0 ratings. A better analysis might be those with 'ratings_count' > 0.

```{r}
#Filter out those with less than 100 reviews
books_filtered <- books_clean %>% 
  filter(ratings_count >= 100)

#max
books_filtered %>% 
  #select(title, average_rating) %>% 
  slice_max(average_rating, n = 10) 

#min
books_filtered %>% 
  #select(title, contains("rating")) %>% #the number of ratings recieved will be relevant to the average rating given
  slice_min(average_rating, n = 10)
```


### QUESTION 2: Most and least popular authors

some authors only have one rating which does not seem representative. Anonymous is multiple people so not interesting in this context. These have been filtered out first.

```{r}
books_clean_authors <- books_clean %>% 
  group_by(authors) %>% 
  filter(ratings_count >= 100, authors != "Anonymous") %>% 
  summarise(average_rating = mean(average_rating), sum(ratings_count))
  #select(authors, contains("rating")) %>% 
books_clean_authors %>%  
slice_max(average_rating, n = 10)

books_clean_authors %>%  
slice_min(average_rating, n = 10)
```

### QUESTION 3: Duplicate book entries?

Some titles appear to be sets despite the individual books also being included
```{r}
books_clean %>% 
  filter(str_detect(title, "Harry Potter")) %>% 
  arrange(title)
```

There are also duplicate titles but they have different author, pages, and publishing information

### QUESTION 4: Longest books / book length

```{r}
books_clean %>% 
  select(title, num_pages, average_rating) %>% 
  slice_max(num_pages, n = 10)

books_clean %>% 
  summarise(mean_num_pages = mean(num_pages, na.rm = TRUE),
            median_num_pages = median(num_pages, na.rm = TRUE),
            max_num_pages = max(num_pages, na.rm = TRUE),
            min_num_pages = min(num_pages, na.rm = TRUE))
```

### QUESTION 5: How much do books vary from the average rating and book length

```{r}
avg_pages <- books_clean %>% 
  summarise(avg_pages = mean(num_pages, na.rm = TRUE)) %>% 
  pull()

avg_rating <- books_clean %>% 
  summarise(avg_rating = mean(average_rating, na.rm = TRUE)) %>% 
  pull()

books_clean %>% 
  select(title, authors, num_pages, average_rating) %>% 
  mutate(avg_pages_diff = num_pages - avg_pages,
         avg_rating_diff = average_rating - avg_rating) %>% 
  arrange(desc(avg_rating_diff))
```

### QUESTION 6: How many books at each rating (rounded)

```{r}
books_clean %>% 
  mutate(rounded_rating = round(average_rating, 0)) %>% 
  group_by(rounded_rating) %>% 
  summarise(count = n())
```

### QUESTION 7: When were the books published?
```{r}
books_clean %>% 
  select(publication_date, title) %>% 
  arrange(desc(publication_date))
```

The books were published between 1900 and 2010

### Thjimen's question: 
```{r}
books_clean %>% 
  filter(ratings_count > 1) %>% 
  select(title, authors, publication_date, average_rating) %>% 
  mutate(new_pub_date = as.Date(publication_date, "%m/%d/%Y"))
```


### QUESTION 8: Number of books in each language
```{r}
books_clean %>% 
  group_by(language_code) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total))
```

There are 27 different language codes, English is given multiple times due to different versions (UK/US/Canada).
English is the most popular langauge for books in this dataset.

