---
title: "R Notebook"
output: html_notebook
---

# Week 6, day 1, Homework

## 1.1 The Titanic
```{r}
library(janitor)
library(gtools)
library(tidyverse)
source("prob.R")

titanic_data <- read_csv("data/titanic.csv")
```

```{r}
# titanic_data %>% 
#   distinct(class)

titanic_p <- titanic_data %>% 
  mutate(crew_passenger = case_when(
    str_detect(class, "[0-9]") ~ "passenger",
    .default = "crew"
  )) %>% 
  select(crew_passenger, survived)

titanic_p
```

```{r}
p_table <- titanic_p %>% 
  tabyl(crew_passenger, survived) %>% 
  adorn_totals(where = c("row", "col")) %>% 
  adorn_percentages(denominator = "all")
p_table
```
p(survived | crew)
```{r}
# p(survived)
p_surv <- 0.32215677

# p(crew)
p_crew <- 0.4032623

# p(survived and crew)
p_surv_and_crew <- 0.09560489

# dependent or independent?
p_surv * p_crew == p_surv_and_crew

# dependent therefore: p(survived | crew) = p(survived and crew) / p(crew)
p_surv_if_crew <- p_surv_and_crew / p_crew
p_surv_if_crew
```

p(survived | passenger)
```{r}
# p(survived)
#p_surv <- 0.32215677

# p(passenger)
p_pass <- 0.5967377

# p(survived and passenger)
p_surv_and_pass <- 0.22655188

# dependent or independent?
p_surv * p_pass == p_surv_and_pass

# dependent therefore: 
# p(survived | passenger) = p(survived and passenger) / p(passenger)
p_surv_if_pass <- p_surv_and_pass / p_pass
```

```{r}
p_surv_if_crew
p_surv_if_pass
```
Passengers were more likely to survive than crew.

## 1.2 Probability

### Question 1.
If I roll a six-sided dice once, what is the probability of getting a number less than 5?

There are 4 numbers less than 5, therefore: 4/6 = 2/3 = *0.6666667*

### Question 2.
How many outcomes are there in the sample space of an experiment in which we toss three dice?
```{r}
6^3

s_three_dice <- rolldie(3)
nrow(s_three_dice)
```

### Question 3.
In an experiment where we roll four dice:

* how many outcomes are in the event where the total of the four dice equals 5, 6 or 7?
* what is the probability of this event?
* what assumptions do we apply in answering part ii?
```{r}
s_four_dice <- rolldie(4)
outcomes_4 <- nrow(s_four_dice)

s_four_dice %>% 
  mutate(total = rowSums(.)) %>% 
  filter(total %in% c(5, 6, 7)) %>% 
  summarise(p = n()/outcomes_4)
```
Assumptions: Every side on every dice is equally likely to be rolled

### Question 4.
Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?
P = N!/(N-r)!
```{r}
1 / factorial(5)
```


### Question 5.
I roll six dice:

* what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right?
```{r}
1 / 6^6
```
```{r}
permutations(6, 6, repeats.allowed = TRUE) %>% 
  as_tibble() %>% 
  summarise(p = 1/n())
```


* what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order?
```{r}
# 1 / factorial(6)
factorial(6) / 6 ^ 6
```

```{r}
options <- permutations(6, 6, repeats.allowed = TRUE) %>% 
  as_tibble() 

options %>% 
  mutate(total = rowSums(.)) %>%
  filter(isin(., c(6, 5, 4, 3, 2, 1), ordered = FALSE)) %>% 
  summarise(p = n()/nrow(options))
  
```


### Question 6.
If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

0.5



