---
title: "Week 6, day 3"
output: html_notebook
---

Homework
```{r}
library(tidyverse)
library(janitor)
library(infer)

ames <- clean_names(read_csv("data/ames.csv"))
```
# Task 2
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot(aes(lot_area)) +
  geom_histogram(colour = "white", bins = 40)
```
There appears to be data at very high lot areas which skewes the histogram. 
Otherwise it is fairly close to normal.

# Task 3
Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.
```{r}
bootstrap_ames <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
bootstrap_ames %>% 
  visualise(bins = 30)
```
# Task 4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution
 
```{r}
ames_ci_95 <- bootstrap_ames %>% 
  get_ci(level = 0.95)

ames_ci_95
```

```{r}
bootstrap_ames %>% 
  visualise(bins = 30) +
  shade_ci(ames_ci_95)
```
# Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence.  
Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above).   
Is it narrower or broader than the 95%CI? Does that make sense?

```{r}
ames_ci_99 <- bootstrap_ames %>% 
  get_ci(level = 0.99)

ames_ci_99
```

```{r}
bootstrap_ames %>% 
  visualise(bins = 30) +
  shade_ci(ames_ci_99)
```
The 99% CI is broader than the 95% CI. This is as expected.

# Task 6.
Calculate the point estimate of the mean(lot_area)
```{r}
ames %>% 
  summarise(mean = mean(lot_area))
```
# Conclusion
**The mean lot area is 10,148 with a 95% CI of [9,803,	10,531]

# Extension
Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? 
```{r}
ames <- ames %>% 
  mutate(built_before_1920 = case_when(
    year_built < 1920 ~ "Yes",
    .default = "No"
  ))

built_before_ames_200 <- ames %>% 
  specify(response = built_before_1920, success = "Yes") %>% 
  generate(reps = 200, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

```{r}
built_before_ames_200 %>% 
  get_ci(level = 0.95)

built_before_ames_200 %>% 
  summarise(prop = mean(stat))
```
6.0% [5.2%, 6.9%]

```{r}
built_before_ames_5000 <- ames %>% 
  specify(response = built_before_1920, success = "Yes") %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

```{r}
built_before_ames_5000 %>% 
  get_ci(level = 0.95)

built_before_ames_5000 %>% 
  summarise(prop = mean(stat))
```
```{r}
built_before_ames_50000 <- ames %>% 
  specify(response = built_before_1920, success = "Yes") %>% 
  generate(reps = 50000, type = "bootstrap") %>% 
  calculate(stat = "prop")
```

```{r}
built_before_ames_50000 %>% 
  get_ci(level = 0.95)

built_before_ames_50000 %>% 
  summarise(prop = mean(stat))
```


